Hierarchical Cluster Analysis for Tax Data
========================================================

# Synopsis 
In this report we aim to find out the proper way to detect outliers in tax data, especially which type of data to use in the analysis. Three different approaches are explored: Hierarchical cluster based only on original data, Hierarchical cluster based only on ratio data and Hierarchical cluster based only on all data. Since no labels in the dataset are avalible, the performance of this analysis is to be evaluated by hand.

# Preprocessing

1.Remove all the "DIV/0!" with "0".

2.Copy and paste spacial with only values to another worksheet.

# Load data into workspace

```{r}
library(xlsx)
library(lattice)

## set word directory
setwd("/Users/xiangjiang/Documents/Tax/data/nashuipinggushuju/Chu Bu Zheng Li Shu Ju/qiyesuodeshuipeibishuju/")

## read data
data <- read.xlsx("08-09peibishuju.xls", sheetName="Sheet2")

## remove irrelevant attributes
data[c("纳税人识别号","所属年度","纳税人名称.1","纳税人识别号.1","所属时期")] <- list(NULL)

## summary of data
summary(data)
```

# Hierarchical cluster based only on original data
Here original data refers to data without ratio.

```{r}
data_orig <- data[2:29]

## computes and returns the distance matrix computed by using the specified distance measure to compute the distances between the rows of a data matrix
d <- dist(as.matrix(data_orig))

## apply hirarchical clustering 
hc <- hclust(d)

pdf( "Hierarchical cluster based only on original data.pdf", width = 50, height = 50 )
## plot the dendrogram
plot(hc, main = "Hierarchical cluster based only on original data")
dev.off()
plot(hc, labels=FALSE, main = "Hierarchical cluster based only on original data")
print("Please find detail in .pdf file")
```

Distribution for different number of clusters are (ranging from 2 to 15 clusters):
```{r}
counts = sapply(2:15,function(ncl)table(cutree(hc,ncl)))
names(counts) = 2:15
counts
```

For each cluster, the members are:
```{r}
## we pass the cutree function the cluster object and the number of clusters we're interested in
## Cuts a tree, e.g., as resulting from hclust, into several groups either by specifying 
## the desired number(s) of groups or the cut height(s).

for(numClusters in c(2:15)){
    print("===============================")
    print("When number of clusters are:")
    print(numClusters)
    mCluster = cutree(hc, numClusters)
    
    for(memCluster in c(1:numClusters)){
        print("members of cluster ")
        print(memCluster)
        
        print(which(mCluster == memCluster))
        #print("Scatter plot for different observations are:")
        #print(splom(~data_orig,groups=mCluster,auto.key=TRUE))
    }   
}
```

# Hierarchical cluster based only on ratio data
Here ratio data refers to data computed with formulars in Excel.

```{r}
data_ratio <- data[30:44]

## computes and returns the distance matrix computed by using the specified distance measure to compute the distances between the rows of a data matrix
d <- dist(as.matrix(data_ratio))

## apply hirarchical clustering 
hc <- hclust(d)

pdf( "Hierarchical cluster based only on ratio data.pdf", width = 50, height = 50 )
## plot the dendrogram
plot(hc, main = "Hierarchical cluster based only on ratio data")
dev.off()
plot(hc, labels=FALSE, main = "Hierarchical cluster based only on ratio data")
print("Please find detail in .pdf file")
```

For each cluster, the members are:
```{r}
## we pass the cutree function the cluster object and the number of clusters we're interested in
## Cuts a tree, e.g., as resulting from hclust, into several groups either by specifying 
## the desired number(s) of groups or the cut height(s).

for(numClusters in c(2:15)){
    print("===============================")
    print("When number of clusters are:")
    print(numClusters)
    mCluster = cutree(hc, numClusters)
    
    for(memCluster in c(1:numClusters)){
        print("members of cluster ")
        print(memCluster)
        
        print(which(mCluster == memCluster))
        #print("Scatter plot for different observations are:")
        #print(splom(~data_ratio,groups=mCluster,auto.key=TRUE))
    }
}

```

# Hierarchical cluster based on all data
Here all data refers to the original data as well as ratio data.
```{r}
data_all <- data[2:44]

## computes and returns the distance matrix computed by using the specified distance measure to compute the distances between the rows of a data matrix
d <- dist(as.matrix(data_all))

## apply hirarchical clustering 
hc <- hclust(d)

pdf( "Hierarchical cluster based on all data.pdf", width = 50, height = 50 )
## plot the dendrogram
plot(hc, main = "Hierarchical cluster based on all data")
dev.off()
plot(hc, labels=FALSE, main = "Hierarchical cluster based on all data")
print("Please find detail in .pdf file")
```

For each cluster, the members are:
```{r}
## we pass the cutree function the cluster object and the number of clusters we're interested in
## Cuts a tree, e.g., as resulting from hclust, into several groups either by specifying 
## the desired number(s) of groups or the cut height(s).

for(numClusters in c(2:15)){
    print("===============================")
    print("When number of clusters are:")
    print(numClusters)
    mCluster = cutree(hc, numClusters)
    
    for(memCluster in c(1:numClusters)){
        print("members of cluster ")
        print(memCluster)
        
        print(which(mCluster == memCluster))
        #print("Scatter plot for different observations are:")
        #print(splom(~data_all,groups=mCluster,auto.key=TRUE))
    }   
}
```
